---
title: "RT-qPCR of pRPS3-mCh-tRPS3mod (2 reps) and pPGK1-mCh-tRPS3mod (2 reps)"
author: "Jamie Auxillos"
date: "06/05/2020"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup,warning=FALSE,message=FALSE,echo=FALSE}

## knitr options for report generation
knitr::opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE,cache=FALSE,
                      results="show",
                      fig.path="figures_pRPS3_pPGK1_tRPS3mod/pRPS3_pPGK1_tRPS3mod_")

library(tidyverse)
library(cowplot)
library(tidyqpcr)

 # set default theme for graphics
theme_set(theme_cowplot(font_size=11) %+replace% 
              theme(panel.border=element_rect(colour = "grey50",
                                            linetype = "solid",size=0.5),
                    strip.background = element_blank()))
```

## Plate plan
```{r label_plates,dependson="plate_functions"}

# list Targets /Probes
TargetID <- c("mCh-7", "URA3-ORF", "RPS3-ORF", "PGK1-ORF")
TargetIDvalues <-factor(rep(TargetID,levels=TargetID))

StrainlevelsR3 <- c("pRPS3-tRPS3_WT","pRPS3-tRPS3_mod0","pRPS3-tRPS3_modA","pRPS3-tRPS3_modB","pRPS3-tRPS3_modC", "pRPS3-tRPS3_modD", "pRPS3-tRPS3_modE","POT1-ccdB")
StrainvaluesR3 <- factor(rep(StrainlevelsR3,each=3),levels=StrainlevelsR3)
StrainlevelsP1 <- c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modA","pPGK1-tRPS3_modB","pPGK1-tRPS3_modC", "pPGK1-tRPS3_modD", "pPGK1-tRPS3_modE","POT1-ccdB")
StrainvaluesP1 <- factor(rep(StrainlevelsP1,each=3),levels=StrainlevelsP1)

BioReplevels_A <- c("BioRep1","BioRep2","BioRep3")
BioRepvalues_A <- factor(rep(BioReplevels_A,times=8),levels=BioReplevels_A)
BioReplevels_B <- c("BioRep4","BioRep5","BioRep6")
BioRepvalues_B <- factor(rep(BioReplevels_B,times=8),levels=BioReplevels_B)

colkeyR3_1 <- tibble(WellC=1:24,
                 Strain=StrainvaluesR3,
                 BioRep=BioRepvalues_A) 
colkeyR3_2 <- tibble(WellC=1:24,
                 Strain=StrainvaluesR3,
                 BioRep=BioRepvalues_B) 
colkeyP1_1 <- tibble(WellC=1:24,
                 Strain=StrainvaluesP1,
                 BioRep=BioRepvalues_A) 
colkeyP1_2 <- tibble(WellC=1:24,
                 Strain=StrainvaluesP1,
                 BioRep=BioRepvalues_B) 

rowkey <- create_rowkey_4in16(TargetID=TargetIDvalues) 

plateplanR3_1 <-     
    label_plate_rowcol(create_blank_plate(WellR = LETTERS[1:16],WellC=1:24),
                       rowkey,colkeyR3_1%>% 
    unite(SampleID,Strain,BioRep,remove=FALSE))

plateplanR3_2 <-     
    label_plate_rowcol(create_blank_plate(WellR = LETTERS[1:16],WellC=1:24),
                       rowkey,colkeyR3_2%>% 
    unite(SampleID,Strain,BioRep,remove=FALSE))
plateplanP1_1 <-     
    label_plate_rowcol(create_blank_plate(WellR = LETTERS[1:16],WellC=1:24),
                       rowkey,colkeyP1_1%>% 
    unite(SampleID,Strain,BioRep,remove=FALSE))
plateplanP1_2 <-     
    label_plate_rowcol(create_blank_plate(WellR = LETTERS[1:16],WellC=1:24),
                       rowkey,colkeyP1_2%>% 
    unite(SampleID,Strain,BioRep,remove=FALSE))
```

## Display plate
```{r display_plates,fig.height=17,fig.width=24,dependson="label_plates"}
display_plate(plateplanR3_1 %>%
    mutate(Strain,BioRep) )

display_plate(plateplanP1_1 %>%
    mutate(Strain,BioRep) )
```


## Load data and attribute sample information
```{r load_plates,dependson="label_plates",results="show"}
# read my plates
platesR3_1 <- read_tsv( "data/JA_20200121-pRPS3-tRPS3mod-n1-ct.txt",skip=1) %>%
    mutate(Well=Pos,Ct=Cp) %>%
    left_join(plateplanR3_1)

platesR3_2 <- read_tsv( "data/JA_20200204-pRPS3-tRPS3mod-n2-ct.txt",skip=1) %>%
    mutate(Well=Pos,Ct=Cp) %>%
    left_join(plateplanR3_2)

platesP1_1 <- read_tsv( "data/JA_20200221_pPGK1-tRPS3mod-n1-ct.txt",skip=1) %>%
    mutate(Well=Pos,Ct=Cp) %>% 
    left_join(plateplanP1_1)

platesP1_2 <- read_tsv( "data/JA_20200311-pPGK1-tRPS3mod-n2-ct.txt",skip=1) %>%
    mutate(Well=Pos,Ct=Cp) %>%
    left_join(plateplanP1_2)

plates <- bind_rows(platesR3_1,platesR3_2,platesP1_1,platesP1_2) 
summary(plates)
```

```{r plot_defvectors}
pRPS3set <- factor(c("pRPS3-tRPS3_WT","pRPS3-tRPS3_mod0","pRPS3-tRPS3_modA",
              "pRPS3-tRPS3_modB","pRPS3-tRPS3_modC","pRPS3-tRPS3_modD",
              "pRPS3-tRPS3_modE"))
pRPS3ATATTC<- c("pRPS3-tRPS3_WT","pRPS3-tRPS3_mod0","pRPS3-tRPS3_modB")
pRPS3linear <- c("pRPS3-tRPS3_WT","pRPS3-tRPS3_mod0","pRPS3-tRPS3_modA",
                 "pRPS3-tRPS3_modD","pRPS3-tRPS3_modE")
pRPS3stability <- c("pRPS3-tRPS3_WT","pRPS3-tRPS3_mod0","pRPS3-tRPS3_modC")



pPGK1set <- c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modA",
              "pPGK1-tRPS3_modB","pPGK1-tRPS3_modC","pPGK1-tRPS3_modD",
              "pPGK1-tRPS3_modE")
pPGK1ATATTC <- c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modB")
pPGK1linear <- c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modA",
                 "pPGK1-tRPS3_modD","pPGK1-tRPS3_modE")
pPGK1stability <- c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modC")

```


## Plot of unnormalized data
```{r plot_unnormalized,dependson="load_plates",fig.height=12,fig.width=8}
ggplot(data=plates %>% filter(Strain %in% pRPS3set)) +
    geom_point(aes(x=TargetID,y=Ct,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(y="Cycle count to threshold",
         title="All biological and technical replicates, 
         \n unnormalized") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    facet_grid(Strain~Type) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.border=element_rect(fill = NA,linetype=1,
                                    colour = "grey50",size=0.5))

ggplot(data=plates %>% filter(Strain %in% pPGK1set)) +
    geom_point(aes(x=TargetID,y=Ct,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(y="Cycle count to threshold",
         title="All biological and technical replicates, 
         \n unnormalized") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    facet_grid(Strain~Type) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.border=element_rect(fill = NA,linetype=1,
                                    colour = "grey50",size=0.5))
```


```{r plot_unnormalizedpot,dependson="load_plates",fig.height=4,fig.width=8}
ggplot(data=plates %>% filter(Strain %in% c("POT1-ccdB"))) +
    geom_point(aes(x=TargetID,y=Ct,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(y="Cycle count to threshold",
         title="All biological and technical replicates, 
         \n unnormalized") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    facet_grid(Strain~Type) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.border=element_rect(fill = NA,linetype=1,
                                    colour = "grey50",size=0.5))
```


## Further analysis of unnormalised data (pPGK1 strains only (WT, mod0, modA, modB) - looking into the Bioreps and Techreps)

A comparison between unnormalised ct values between pPGK1 strains was carried out to identify if the high  Ct values for pPGK1-tRPS3_mod0 had occurred in one biorep and if it is present in all techreps for this one sample.

Conclusion - all ct values in the +RT column higher than 20 were only present in the pPGK1-tRPS3_WT Strain. In particular, the pPGK1-tRPS3_WT Biorep 2 sample. High Ct values were observed for all 3 technical replicates for all 4 primers (TargetID).

**Take note, pPGK1-tRPS3_WT Biorep 1, with PGK1-ORF seems to have one techrep (1) with a lower Ct value. pPGK1-tRPS3_WT Biorep 1, with RPS3-ORF seems to have one techrep (1) with a higher Ct value. 
```{r plot_unnormalized_pPGK1analysis,dependson="load_plates",fig.height=8,fig.width=8}
ggplot(data=plates %>% filter(Strain %in% c("pPGK1-tRPS3_WT","pPGK1-tRPS3_mod0","pPGK1-tRPS3_modA","pPGK1-tRPS3_modB"))) +
    geom_point(aes(x=TargetID,y=Ct,colour=BioRep,shape=TechRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(y="Cycle count to threshold",
         title="All reps, unnormalized, 
         \n Analysis of pPGK1 data points") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    facet_grid(Strain~Type) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.border=element_rect(fill = NA,linetype=1,
                                    colour = "grey50",size=0.5))

pPGK1data<- (data=plates %>% filter(SampleID %in% c("pPGK1-tRPS3_WT_BioRep1","pPGK1-tRPS3_WT_BioRep2","pPGK1-tRPS3_WT_BioRep3")))
pPGK1data

```

```{r plot_unnormalized_BiorepWTanalysis,dependson="load_plates",fig.height=4,fig.width=8}
ggplot(data=plates %>% filter(SampleID %in% c("pPGK1-tRPS3_WT_BioRep1","pPGK1-tRPS3_WT_BioRep2","pPGK1-tRPS3_WT_BioRep3","pPGK1-tRPS3_WT_BioRep4","pPGK1-tRPS3_WT_BioRep5","pPGK1-tRPS3_WT_BioRep6"))) +
    geom_point(aes(x=TargetID,y=Ct,colour=BioRep,shape=TechRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(y="Cycle count to threshold",
         title="All reps, unnormalized, 
         \n pPGK1-tRPS3_WT_BioRep comparison") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    facet_grid(Strain~Type) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.border=element_rect(fill = NA,linetype=1,
                                    colour = "grey50",size=0.5))
```

## Normalisation of mCherry Ct values against PGK1-ORF Ct values
Here, we are excluding pPGK1-tRPS3_WT_BioRep2 data because it seemed to show a problem in the sample preparation (high Ct value for all technical replicates for this one biological replicate)
```{r normalize_counts,dependson="load_plates"}
platesnorm <- plates  %>% filter(!SampleID %in% c("pPGK1-tRPS3_WT_BioRep2")) %>%
    filter(Type=="+RT") %>%
    normalizeqPCR(normTargetIDs = "PGK1-ORF") 
    #%>%filter(TargetID=="mCh-7")

platesmed <- platesnorm %>% filter(!SampleID %in% c("pPGK1-tRPS3_WT_BioRep2")) %>%
      group_by(Strain,TargetID,BioRep) %>%
     # filter(TargetID=="mCh-7")%>%
      summarize(Ct=median(Value.norm,na.rm=TRUE),
              Abund=median(Value.normexp,na.rm=TRUE))
      filter(platesmed,TargetID=="PGK1-ORF")
```

## Plotting all the normalised (to PGK1-ORF) Ct values for all biological replicates and technical replicates for each strain
```{r plot_diffnormalization,dependson="normalize_counts",fig.height=5,fig.width=6}
platesnorm


ggplot(data=platesnorm %>% filter (TargetID=="mCh-7")) +
    geom_point(aes(x=Value.norm,y=Strain, colour=BioRep, shape=TechRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(x="Ct relative to PGK1-ORF",
    title="Normalised mCherry Ct values  
         \n relative to PGK1-ORF") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    xlim(-5, 15)+
    facet_grid(TargetID~.) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.grid.major = element_line(colour = "grey50",size=0.125))


ggplot(data=platesnorm %>% filter (TargetID=="RPS3-ORF")) +
    geom_point(aes(x=Value.norm,y=Strain, colour=BioRep, shape=TechRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(x="Ct relative to PGK1-ORF",
    title="Normalised RPS3-ORF Ct values  
         \n relative to PGK1-ORF") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    xlim(-5, 15)+
    facet_grid(TargetID~.) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.grid.major = element_line(colour = "grey50",size=0.125))

ggplot(data=platesnorm %>% filter (TargetID=="URA3-ORF")) +
    geom_point(aes(x=Value.norm,y=Strain, colour=BioRep, shape=TechRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(x="Ct relative to PGK1-ORF",
    title="Normalised URA3-ORF Ct values  
         \n relative to PGK1-ORF") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    xlim(-5, 15)+
    facet_grid(TargetID~.) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          panel.grid.major = element_line(colour = "grey50",size=0.125))
```


## Plotting the mCherry mRNA detected in each strain (for biological replicates only)
```{r plot_pRPS3_mChnormalized,dependson="normalize_counts",fig.height=3,fig.width=6}

platesmed$Strain <- factor(platesmed$Strain)

ggplot(data=platesmed %>% filter(Strain %in% pRPS3set) %>% 
         filter (TargetID=="mCh-7")) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="All normalised mCherry to PGK1-ORF 
         \n of pRPS3-mCherry strains") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(0.001, 8)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))


platesmed_pRPS3ATATTC <- platesmed %>% filter(Strain %in% pRPS3ATATTC) %>% filter (TargetID=="mCh-7")
platesmed_pRPS3ATATTC$Strain <- factor(platesmed_pRPS3ATATTC$Strain,
                                       levels=c("pRPS3-tRPS3_modB","pRPS3-tRPS3_mod0", "pRPS3-tRPS3_WT"))

ggplot(data=platesmed_pRPS3ATATTC) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
    labs(title="Normalised mCherry to PGK1-ORF of 
         \n pRPS3-mCherry ATATTC construct") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(0.001, 8)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))




platesmed_pRPS3linear <- platesmed %>% filter(Strain %in% pRPS3linear) %>% filter (TargetID=="mCh-7")
platesmed_pRPS3linear$Strain <- factor(platesmed_pRPS3linear$Strain,
                                       levels=c("pRPS3-tRPS3_modE","pRPS3-tRPS3_modD","pRPS3-tRPS3_modA",
                                                "pRPS3-tRPS3_mod0", "pRPS3-tRPS3_WT"))
ggplot(data=platesmed_pRPS3linear) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pRPS3-mCherry linear combination of decay motif constructs") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(0.001, 8)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))



platesmed_pRPS3stability <- platesmed %>% filter(Strain %in% pRPS3stability) %>% filter (TargetID=="mCh-7")
platesmed_pRPS3stability$Strain <- factor(platesmed_pRPS3stability$Strain,
                                       levels=c("pRPS3-tRPS3_modC","pRPS3-tRPS3_mod0", "pRPS3-tRPS3_WT"))
ggplot(data=platesmed_pRPS3stability) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pRPS3-mCherry stability motif construct") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(0.001, 8)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))
```



```{r plot_pPGK1_mChnormalized,dependson="normalize_counts",fig.height=3,fig.width=6}

ggplot(data=platesmed %>% filter(Strain %in% pPGK1set) %>% 
         filter (TargetID=="mCh-7")) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pPGK1-mCherry strains") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(-5, 15)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))


platesmed_pPGK1ATATTC <- platesmed %>% filter(Strain %in% pPGK1ATATTC) %>% filter (TargetID=="mCh-7")
platesmed_pPGK1ATATTC$Strain <- factor(platesmed_pPGK1ATATTC$Strain,
                                       levels=c("pPGK1-tRPS3_modB","pPGK1-tRPS3_mod0", "pPGK1-tRPS3_WT"))
ggplot(data=platesmed_pPGK1ATATTC) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pPGK1-mCherry ATATTC construct") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(-5, 15)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))



platesmed_pPGK1linear <- platesmed %>% filter(Strain %in% pPGK1linear) %>% filter (TargetID=="mCh-7")
platesmed_pPGK1linear$Strain <- factor(platesmed_pPGK1linear$Strain,
                                       levels=c("pPGK1-tRPS3_modE","pPGK1-tRPS3_modD","pPGK1-tRPS3_modA",
                                                "pPGK1-tRPS3_mod0", "pPGK1-tRPS3_WT"))
ggplot(data=platesmed_pPGK1linear) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pPGK1-mCherry linear combination of decay motif constructs") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(-5, 15)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))




platesmed_pPGK1stability <- platesmed %>% filter(Strain %in% pPGK1stability) %>% filter (TargetID=="mCh-7")
platesmed_pPGK1stability$Strain <- factor(platesmed_pPGK1stability$Strain,
                                       levels=c("pPGK1-tRPS3_modC","pPGK1-tRPS3_mod0", "pPGK1-tRPS3_WT"))
ggplot(data=platesmed_pPGK1stability) +
    geom_point(aes(x=Abund,y=Strain,colour=BioRep),
               position=position_jitter(width = 0.2,height=0)) +
      labs(x="Abund",
         title="Normalised mCherry to PGK1-ORF of 
         \n pPGK1-mCherry stability motif construct") +
    scale_shape_manual(values=c(15:18,5:6)) + 
    scale_x_log10nice("Relative mRNA detected") + 
    xlim(-5, 15)+
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
         panel.grid.major = element_line(colour = "grey50",size=0.125))
```

